### 2013-10-29 src2/thesis.tex
update

\subsection{Writing efficient GPU kernels}

@@

\subsection{Results}

@@

% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
%
% ArBB work
%
% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\section{Embedding Intel Array Building Blocks}

@@

\section[\paperFTitle]{\paperF: \\ \paperFTitle}
\label{sec:paperF}
%\addcontentsline{toc}{chapter}{Simple and Compositional Reification of Monadic Embedded Languages}

@@

>>>>

\subsection{Writing efficient GPU kernels} 
\label{sec:efficient}

@@

\subsection{Results}
\FloatBarrier

@@

In section \ref{sec:efficient}, guidelines for writing efficient CUDA code is listed. 
Here we revisit these guidelines and explain how they are approached by Obsidian and 
the Obsidian programmer.

\item {\bf Minimise data transfers between host and device:} Data transfers
between host and device are out of Obsidian's scope. If Obsidian is used to create 
a kernel for a computation, the programmer has already decided that a GPU should be used.

\item {\bf Ensure that global memory accesses are coalesced:} Inputs to an Obsidian 
program are pull arrays. Pull arrays can be arbitrarily permuted; the permutation 
used will decide the access pattern. If this method can be applied or not partly also 
depends on the algorithm being implemented.

\item {\bf Minimise use of global memory:} Intermediate results can be computed into 
shared memory. Arrays are stored into shared memory when the programmer uses a 
{\tt force} primitive.

\item {\bf Use a multiple of 32 threads per block:} Obsidian requires regularity even more 
than CUDA does. Since we focus on generating high performance kernels, we rule out generation 
of kernels that work on different input sizes. The choise of how many threads to use per block 
is in the hands of the Obsidian program. The method for making that choise has changed over time. 
Earlier, the number of threads needed by kernel was decided by the size of the arrays it 
computed. This has recently changed is now set by parameters to the code generation process.

\item {\bf Avoid different execution paths within a block:} Sometimes divergent code cannot 
be avoided. However, It is important not to have threads diverge unnecessarily. When using pull 
arrays alone this was often a problem; with the addition of push arrays the situation improved. 
Using push arrays a divergent computation can be explicitly split up into phases, each with all 
threads following the same execution path. This means that fewer threads are required to do the 
same work.

\item {\bf Do not use {\tt \_\_synchthreads()} in diverging code:} In Obsidian it is impossible 
for a barrier synchronisation to appear within conditional code. Conditionals can be introduced 
either by a conditional expression or a {\tt Cond} operation in the program datatype. The Cond 
operation chooses between two thread programs and the synchronisation operation is at the block 
level in the hierarchy. So a synchronisation cannot occur in conditional code. 
%f it is safe or 
%not to call some function at a certain place in your code, can only be decided by inspecting 
%that functions implementation. 
\end{itemize}

%% In the book ``Structured Parallel Programming''~\citet{STRUCTURED}, there is a list showing language 
%% support for various parallel patterns. In figure~\ref{fig:caps} this list augmented with a column for Obsidian. 
%% This figure is added to indicate relative expressive power of the different languages. The most 
%% interesting comparison is between Obsidian and OpenCL that both have a programming model geared towards
%% GPUs; even though OpenCL can compile for both CPU and GPU execution. The paragraphs below 
%% take a closer look at well Obsidian performes in the patterns.

%% \subsubsection{Map}

%% Map, to a  functional programmer is a higher order function that takes a pure 
%% function and applies it to all elements of a collection (for example a list). 
%% That meaning of map is easily implementable in Obsidian on both pull and push arrays. 
%% Obsidian is also capable, via its parallel for loops, to map sequential computations 
%% over threads in a block and block computation over many blocks. This is probably 
%% more in the spirit of what the authors refer to in~\citet{STRUCTURED}.

%% \subsubsection{Stencil}

%% Stencil computations can be expressed in Obsidian, with some manual labour. 
%% A stencil computation would have the form of a pull array to pull array function. 
%% Pull arrays allow reading at position. 
%% Stencil operations where more than one output position is touched could be 
%% implemented as a pull array to push array function.
%% One problem with stencil computations is the borders. If every thread needs to 
%% execute a conditional, checking if it is at the border, performance would suffer. 
%% Using Obsidian the programmer would likely generate more than one kernel; a fast 
%% kernel for the internal points and one or more border kernel.

%% \subsubsection{Reduction and scan} 
%% Reductions can easily be implemented in Obsidian and with very good performance. 
%% Obsidian also seems particularly well suited for experimenting with the kind 
%% of trade offs that are interesting when implementing reduction kernels.

%% Scans can also be implemented and the code is elegant.

%% \subsubsection{Pack and expand}

%% \begin{figure}[!ht]
%% \begin{small} 
%% \begin{tabular}{|l|l|l|l|l|l|l|}
%% Parallel Pattern & TBB & Cilk Plus & OpenMP & ArBB & OpenCL & Obsidian \\
%% \hline 
%% Parallel Nesting & F   & F         &        &      &        &          \\
%% Map              & F   & F         & F      & F    & F      & I        \\
%% Stencil          & I   & I         & I      & F    & I      & I        \\ 
%% Workpile         & F   &           &        &      & I      &          \\
%% Reduction        & F   & F         & F      & F    & I      & I        \\ 
%% Scan             & F   & I         & I      & F    & I      & I        \\ 
%% Fork-Join        & F   & F         & I      &      &        &          \\ 
%% Recurrance       &     & P         &        &      &        &          \\ 
%% Superscalar seq. &     &           &        &      & F      &          \\

%% Pack             & I   &  I        &  I     & F    & I      & I        \\ 
%% Expand           & I   &  I        &  I     & I    & I      & *        \\
%% Pipeline         & F   &  I        &  I     &      &        &          \\
%% Geometric decomp. & I  &  I        &  I     & I    & I      & I        \\
%% Search           &  I  &  I        &  I     & I    & I      & I        \\
%% Category reduction & I &  I        &  I     & I    & I      & I        \\
%% Gather           & I   &  F        &  I     & F    & I      & I        \\
%% Atomic scatter   & F   &  I        &  I     &      & I      & I        \\
%% Permutation scatter & F & F        &  F     & F    & F      & I        \\ 
%% Merge scatter    &  I  &  I        &  I     & F    & I      & *        \\
%% Priority scatter &     &           &        &      &        & *        \\

%% \end{tabular}
%% \end{small}
%% \label{fig:caps}
%% \caption{This figure lists what parallel patterns a number of languages support. The list 
%% is copied from ``Structured Parallel Programming''~{\protect\citet{STRUCTURED}} and augmented with a 
%% column for Obsidian. \\
%% {\bf Key:} F - means the pattern is supported directly with an abstraction. P - indicates 
%% that the pattern can be implemented in terms of other patterns appearing below. I - means 
%% that the pattern can be easily implemented. Lastly, (*) indicates some patterns 
%% that can be implemented in Obsidian but currently not efficiently so.} 
%% \end{figure}

%% In the technical report ``The Landscape of Parallel Computing Research: A View 
%% from Berkeley''~\citet{dwarfs}, 13 computational kernels (called dwarfs) of importance is identified. 
%% The paper also proposes that we use these 13 dwarfs to evaluate parallel programming 
%% models and languages. The dwarfs are listed below, together with thoughts on how 
%% well they can be tackled by Obsdian.

%% \begin{itemize} 
%% \item {\bf Dense Linear Algebra:} Obsidian works well for expressing vector-vector 
%% operations. It is also possible to express operations such as matrix multiplication, 
%% even though Obsidians support for multidimensional arrays is limited. It is possible 
%% to use pull arrays of pull arrays, but not push arrays of push arrays.

%% \item {\bf Sparse Linear Algebra:}

%% \item {\bf Spectral Methods (FFT):} FFTs are implemented in a way that is 
%% similar to sorting networks. A so-called butterfly network is used. We know 
%% that we can express these network structures in Obsidian.

%% \item {\bf N-Body Methods:}

%% \item {\bf Structured Grids:} Regular grids should not be a problem for Obsidian. 
%% Obsidian has no built in support for stencils, the combination of pull and push array 
%% still provides the capability to express such computations.

%% \item {\bf Unstructured Grids:} May be hard problem for GPUs in general. The paper 
%% states that these problems have the property that ``Updates typically involve multiple 
%% levels of memory reference indirection''. We know that the GPU thrives when memory access 
%% patters are predictable and regular, which is the opposite scenario.

%% \end{itemize}

\FloatBarrier 
% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
%
% ArBB work
%
% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\section{Embedding Intel Array Building Blocks}

@@

\section[\paperFTitle]{\paperF: \\ \paperFTitle}
\label{sec:paperF}

%\addcontentsline{toc}{chapter}{Simple and Compositional Reification of Monadic Embedded Languages}

@@

\begin{center}
{\large \bf This paper will be revised before printing the Thesis} 
\end{center}
